Design and Implementation of Probabilistic Programming Language Anglican

### 2016-08-04 paper/paper.tex
update

    ;; Predicting a, b and the distribution
    (predict :a a) 
    (predict :b b)
    (predict :d d)))
\end{lstlisting}
This query defines a probability distribution on predicted
variables, $\texttt{d} \in \{\texttt{normal},\texttt{gamma}\}$
for a distribution type, and $\texttt{a}$ and $\texttt{b}$ for
positive parameters for the type. Samples from this distribution 
can be obtained by running one of Anglican's inference algorithms
on the query.

@@

Anglican is intended to co-exist with Clojure and be a part of
the source of a Clojure program. To facilitate this, Anglican
programs, or queries, are wrapped by macros (defined in the
\texttt{anglican.emit} namespace), which call the CPS
transformations and define Clojure objects suitable for passing
as arguments to inference algorithms (\texttt{defquery},
\texttt{query}). In addition to defining entire queries,
Anglican promotes modularization of inference algorithms through
definition of \textit{probabilistic functions} using
\texttt{defm} and \texttt{fm} (Anglican counterparts of Clojure
\texttt{defn} and \texttt{fn}). Probabilistic functions are
written in Anglican, may include probabilistic forms
\texttt{sample} and \texttt{observe} (as well \texttt{predict}
for the output), and can be seamlessly called from inside
Anglican queries, just like functions locally defined within the
same query.

Operational semantics of Anglican queries is different from that
of Clojure code, therefore queries must be called through
inference algorithms, rather than `directly'.  The
\texttt{anglican.inference} namespace supplies the
\texttt{infer} multimethod, which accepts an Anglican query and
returns a lazy sequence of weighted samples from the
distribution defined by the query.  When inference
is performed on an Anglican query, the query is run by a
particular inference algorithm. Inference algorithms must
provide an implementation for \texttt{infer}, as well as
override some of the methods of the \texttt{checkpoint}
multimethod, called to handle \texttt{sample} and
\texttt{observe} in an algorithmic-specific manner, as well
as on termination of a probabilistic program.

Finally, Anglican queries use `primitive', or commonly known
and used, distributions, to draw random samples and condition
observations. Many primitive distributions are provided by the
\texttt{anglican.runtime} namespace, and additional
distributions can be defined by the user by implementing the
\texttt{distribution} protocol. The \texttt{defdist} macro
provides a convenient syntax for defining primitive distributions.

>>>>

    ;; Predicting a, b and the distribution
    (predict :a a) 
    (predict :b b)
    (predict :d d)))
\end{lstlisting}
This query defines a probability distribution on three predicted
variables, $\texttt{d} \in \{\texttt{normal},\texttt{gamma}\}$
for a distribution type, and $\texttt{a}$ and $\texttt{b}$ for
positive parameters for the type. Samples from this distribution 
can be obtained by running the query under one of Anglican's 
inference algorithms.

@@

Anglican is intended to co-exist with Clojure and be a part of
the source of a Clojure program. To facilitate this, Anglican
programs, or queries, are wrapped by macros (defined in the
\texttt{anglican.emit} namespace), which call the CPS
transformations and define Clojure values
%Clojure objects
suitable for passing
as arguments to inference algorithms (\texttt{defquery},
\texttt{query}). In addition to defining entire queries,
Anglican promotes modularization of inference algorithms through
the definitions of \textit{probabilistic functions} using
\texttt{defm} and \texttt{fm} (Anglican counterparts of Clojure
\texttt{defn} and \texttt{fn}). Probabilistic functions are
written in Anglican, may include probabilistic forms
\texttt{sample} and \texttt{observe} (as well \texttt{predict}
for the output), and can be seamlessly called from inside
Anglican queries, just like functions locally defined within the
same query.

Operational semantics of Anglican queries is different from that
of Clojure code, therefore queries must be called through
inference algorithms, rather than `directly'.  The
\texttt{anglican.inference} namespace declares the polymorphic
function \texttt{infer} using Clojure's multimethod mechanism.
This function accepts an Anglican query and
returns a lazy sequence of weighted samples from the
distribution defined by the query.  When inference
is performed on an Anglican query, the query is run by a
particular inference algorithm. Inference algorithms must
provide an implementation for \texttt{infer}, as well as
override the polymorphic function \texttt{checkpoint} (defined
as a multimethod) so as to handle \texttt{sample} and
\texttt{observe} in an algorithm-specific manner, as well
as to construct an appropriate result on the 
termination of a probabilistic program.

Finally, Anglican queries use `primitive', or commonly known
and used, distributions, to draw random samples and condition
observations. Many primitive distributions are provided by the
\texttt{anglican.runtime} namespace, and an additional
distribution can be defined by the user by implementing a
particular set of functions for the distribution 
(via Clojure's protocol mechanism). The \texttt{defdist} macro
provides a convenient syntax for defining primitive distributions.

\hy{I think that the concepts like multimethod and protocol
are specific to Clojure, and they are not well-known among PL
people. I tried to rephase the text so that people can follow
it even when they don't know these concepts. I didn't know
``form'' before, but using it is perhaps ok because it is a
standard terminology for Lisp programmers.}
